---
title: "R Assignment #1"
author: "[Kleber Perez - `r format(Sys.Date(), '%B %d, %Y')`](https://www.linkedin.com/in/kleberperez/)" 
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(ggplot2)
```


# <span style="color:#9f0a0a">Mushrooms Data</span>

<br />

> The Mushrooms data is a well-known dataset in the data science community makes it a good dataset to use for comparative benchmarking.  For example, if someone was working to build a better decision tree algorithm (or other predictive classifier) to analyze categorical data, this dataset could be useful. (https://archive.ics.uci.edu/ml/datasets/Mushroom)

> *Display original dataset columns.*

```{r}
mushrooms <- read_csv("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")
```

<br />

> **1. Use the summary function to gain an overview of the data set. Then display attributes.**

```{r}
summary(mushrooms)
```

<br />

## <span style="color:#9f0a0a">Subsetting Data</span>

<br />

> **2. Create a new data frame with a subset of the columns**

```{r}
subsetMushrooms <- mushrooms[c(1, 2, 3:4)]
```

<br />

> **3. Create new column names for the new data frame.**

```{r}
colnames(subsetMushrooms) <- c("class", "bruises", "gill-size", "habitat")
DT::datatable(subsetMushrooms, options = list(pageLength = 10))
```

<br />

> **4. Use the summary function to create an overview of your new data frame and print attributes.**

```{r}
summary(subsetMushrooms)
```

<br />

> **5. 4.	Replace abbreviation with meaningful names (i.e.;  e = edible)**

```{r}
#newdata <- subset(SubsetMush, class >= "e" & Class <= "p")
newdata <- subset(subsetMushrooms, class == "e" | class <= "p")
newdata [newdata == "e"] <- "Edible"
newdata [newdata == "p"] <- "Poisonous"
```

<br />


> **6. See examples of all of steps 1-5 above.**

```{r}
DT::datatable(newdata, options = list(pageLength = 10))  
```

<br />

## <span style="color:#9f0a0a">Plotting </span>


```{r}
library(tidyverse)
#library(ggplot2)
#library(dplyr)
#factorMushrooms <- mushrooms[c(1, 3, 9, 23)]
#factorMushrooms <- read_csv("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")

subsetMushrooms <- mushrooms[c(1, 3, 9, 23)]
colnames(subsetMushrooms) <- c("class", "capsurface", "gillsize", "habitat")

summary(subsetMushrooms)

subsetMushrooms <- subsetMushrooms %>%
  mutate(class=as.factor(class), capsurface=as.factor(capsurface), 
         gillsize=as.factor(gillsize), habitat=as.factor(habitat))
  #mutate(class=factor(class), habitat=factor(habitat))
#unique(factorMushrooms$class)

summary(subsetMushrooms)
levels(subsetMushrooms$class)
levels(subsetMushrooms$capsurface)
levels(subsetMushrooms$gillsize)
levels(subsetMushrooms$habitat)

# Set up factors
#df <- data.frame(a = factorMushrooms$class, factorMushrooms$bruises)

ggplot(data = subsetMushrooms) + 
  geom_point(mapping = aes(x=class, y=habitat))

```


```{r}
#subsetMushrooms <- subsetMushrooms[c(1, 3, 9, 23)]

require(data.table) 
counts<-data.frame(table(subsetMushrooms))
DT::datatable(counts, options = list(pageLength = 10))  

plot(subsetMushrooms)
plot(counts$class,counts$capsurface)

```


<br />

> *Please email to: * <a href="mailto:kleber.perez@live.com">kleber.perez@live.com</a> *for any suggestion.*

<br />
<br />
<br />
<br />

<footer class="border-top text-muted fixed-bottom">

        ** R Assignment #1 - MSDS 2019 Program.

</footer>

\textbf{Text ignored: END of other links}



